Body size patterns

### 2019-03-22 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
added ...bySpp.Rmd

---
title: "Growth model analysis 3spp WB"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

@@

```{r von Bert, well, gam}

@@

vB$riverFactor <- factor(vB$river, levels = c('bkt','bnt'))
#vB$riverOrdered <- riverOrdered exists

@@

################################################################################

@@

mod <- bam(observedLength ~ cohort * river + 
               s(ydayCumul, bs = 'cr'), 
               data = vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt')), 
               method = "REML")
summary(mod)

@@

##########################################################
# by cohortSpeciesRiver 
kIn <- 4
system.time(
modCR <- bam(observedLength ~ cohort * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered), 
             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt')), method = "REML")
)
summary(modCR)
plot(modCR, pages = 8, scale = 0)

@@

predGrid2CR <- bind_cols(predGridCR,as.data.frame(predCR))

ggplot(predGrid2CR, aes(ydayCumul,predCR)) +
  geom_point() + 
  geom_line() +
  ylim(50,210) +
  facet_grid(river ~ cohort)+
  ggtitle('modCR')

##########################################################
# by cohortSpeciesRiver, additive in s() 
kIn <- 4 # 4 looks good
system.time(
modC_R <- bam(observedLength ~ cohort * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered), 
             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt')), method = "REML")
)
summary(modC_R)
plot(modC_R, pages = 4, scale = 0)

@@

##########################################################
# by cohortSpeciesRiver - additive splines, no 3-way spline
system.time(
modC_S_R <- bam(observedLength ~ cohort * species * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', by = speciesOrdered) +
               s(ydayCumul, bs = 'cr', by = riverOrdered) +
               s(ydayCumul, bs = 'cr', by = cohortRiverOrdered) +
               s(ydayCumul, bs = 'cr', by = speciesCohortOrdered) +
               s(ydayCumul, bs = 'cr', by = speciesRiverOrdered), 
             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
)
summary(modC_S_R)
#plot(modC_S_R)

predGridC_S_R <- expand.grid(cohortSpeciesRiverOrdered = unique(modCSR$model$cohortSpeciesRiverOrdered), # use modCSR to get 3-way
                               ydayCumul = seq(250,1000, by = 30)) %>%
                    mutate(cohortOrdered = as.numeric(substr(cohortSpeciesRiverOrdered,1,4)),
                           speciesOrdered = substr(cohortSpeciesRiverOrdered,6,8),
                           riverOrdered = substring(cohortSpeciesRiverOrdered,10),
                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
                           speciesCohortOrdered = paste(speciesOrdered, cohortOrdered, sep = '_'),
                           speciesRiverOrdered = paste(speciesOrdered, riverOrdered, sep = '_'),
                           cohortCentered = cohortOrdered - cohortOffset,
                           cohort = cohortOrdered,
                           species = speciesOrdered,
                           river = riverOrdered)

predC_S_R <- predict(modC_S_R, newdata = predGridC_S_R)#, type = 'lpmatrix')

predGrid2C_S_R <- bind_cols(predGridC_S_R,as.data.frame(predC_S_R))

ggplot(predGrid2C_S_R, aes(ydayCumul,predC_S_R, color = species)) +
  geom_point() + 
  geom_line() +
  ylim(50,210) +
  facet_grid(river ~ cohort) +
  ggtitle('modC_S_R')

##########################################################
# by speciesRiver

modSR <- bam(observedLength ~ cohort * species * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', by = speciesOrdered) +
               s(ydayCumul, bs = 'cr', by = riverOrdered) +
               s(ydayCumul, bs = 'cr', by = speciesRiverOrdered), 
             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
summary(modSR)
plot(modCR, scale = 0)

predGridSR <- expand.grid(speciesRiverOrdered = unique(modSR$model$speciesRiverOrdered),
                          cohortOrdered = 2000:2013,
                          ydayCumul = seq(250,1000, by = 30)) %>%
                   mutate(speciesOrdered = substr(speciesRiverOrdered,1,3),
                          riverOrdered = substring(speciesRiverOrdered,5),
                          cohortCentered = cohortOrdered - cohortOffset,
                          cohort = cohortOrdered,
                          species = speciesOrdered,
                          river = riverOrdered)

predSR <- predict(modSR, newdata = predGridSR)#, type = 'lpmatrix')

predGrid2SR <- bind_cols(predGridSR,as.data.frame(predSR))

ggplot(predGrid2SR, aes(ydayCumul,predSR, color = species)) +
  geom_point() + 
  geom_line() +
  facet_grid(river ~ cohortOrdered)+
  ggtitle('modSR')

#######################################
# by speciesCohort
system.time(
modSC <- bam(observedLength ~ cohort * species * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', by = speciesOrdered) +
               s(ydayCumul, bs = 'cr', by = speciesCohortOrdered), 
             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
)
summary(modSC)
#plot(modSC, scale = 0)

predGridSC <- expand.grid(speciesCohortOrdered = unique(modSC$model$speciesCohortOrdered),
                          river = unique(vB$riverOrdered),
                          ydayCumul = seq(250,1000, by = 30)) %>%
                   mutate(speciesOrdered = substr(speciesCohortOrdered,1,3),
                          cohortOrdered = as.numeric(substring(speciesCohortOrdered,5)),
                          cohortCentered = cohortOrdered - cohortOffset,
                          cohort = cohortOrdered,
                          species = speciesOrdered)

predSC <- predict(modSC, newdata = predGridSC)#, type = 'lpmatrix')

predGrid2SC <- bind_cols(predGridSC,as.data.frame(predSC))

ggplot(predGrid2SC, aes(ydayCumul,predSC, color = species)) +
  geom_point() + 
  geom_line() +
  facet_grid(river ~ cohortOrdered) +
  ggtitle('modSC')

##################################
# by cohortRiver

modCR <- bam(observedLength ~ cohort * species * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', by = riverOrdered) +
               s(ydayCumul, bs = 'cr', by = cohortRiverOrdered), 
               data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
summary(modCR)
#plot(modCR, scale = 0)

predGridCR <- expand.grid(cohortRiverOrdered = unique(modCR$model$cohortRiverOrdered),
                          speciesOrdered = c('bkt','bnt'),
                          ydayCumul = seq(250,1000, by = 30)) %>%
                   mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
                          riverOrdered = (substring(cohortRiverOrdered,6)),
                          cohort = cohortOrdered,
                          cohortCentered = cohortOrdered - cohortOffset,
                          river = riverOrdered,
                          species = speciesOrdered)

ggplot(predGrid2CR, aes(ydayCumul, predCR, color = species)) +
  geom_point() + 
  geom_line() +
  ylim(50,200) +
  facet_grid(river ~ cohortOrdered) +
  ggtitle('modCR')

#############################

# add year to smooths, model seems overfit

# try running without 'speciesOrdered' for year  this does a difference from the main fit, i.e. try s(year, bs = 'cr', by = speciesFactor)

modCSRY <- bam(observedLength ~ cohort * species * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', by = speciesOrdered) +
               s(ydayCumul, bs = 'cr', by = riverOrdered) +
               s(ydayCumul, bs = 'cr', by = cohortRiverOrdered) +
               s(ydayCumul, bs = 'cr', by = speciesCohortOrdered) +
               s(ydayCumul, bs = 'cr', by = speciesRiverOrdered) +
               s(ydayCumul, bs = 'cr', by = cohortSpeciesRiverOrdered) +
               #s(year, bs = 'cr', by = speciesOrdered), 
               s(year, bs = 'cr', by = speciesFactor),
               data = vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt','bnt')), 
               method = "REML")
summary(modCSRY)
#plot(modCSRY, select = 189)

predGridCSRY <- expand.grid(cohortSpeciesRiverOrdered = unique(modCSRY$model$cohortSpeciesRiverOrdered),
                            year = unique(modCSRY$model$year),
                            ydayCumul = seq(250,1000, by = 30)) %>%
                    mutate(cohortOrdered = as.numeric(substr(cohortSpeciesRiverOrdered,1,4)),
                           speciesOrdered = substr(cohortSpeciesRiverOrdered,6,8),
                           riverOrdered = substring(cohortSpeciesRiverOrdered,10),
                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
                           speciesCohortOrdered = paste(speciesOrdered, cohortOrdered, sep = '_'),
                           speciesRiverOrdered = paste(speciesOrdered, riverOrdered, sep = '_'),
                           cohortCentered = cohortOrdered - cohortOffset,
                           cohort = cohortOrdered,
                           species = speciesOrdered,
                           river = riverOrdered)

predCSRY <- predict(modCSRY, newdata = predGridCSRY)#, type = 'lpmatrix')

predGrid2CSRY <- bind_cols(predGridCSRY,as.data.frame(predCSRY))

ggplot(predGrid2CSRY, aes(ydayCumul, predCSRY, color = species)) +
  geom_point() + 
  geom_line() +
  ylim(50,300) +
  facet_grid(river ~ cohortOrdered) +
  ggtitle('modCSRY')

@@

save(mod,modCR,modCSR,modSC,modSR,modCSRY,modC_S_R, file = './data/out/mods.RData')
AIC(modC_S_R,modCSR,modSR,modSC,modCR,mod)  %>% rownames_to_column() %>% arrange(AIC)

@@

```{r mean lengths by grouping Variable}

#####################################################
#####################################################

# Select THIS for whole pop analysis, or
groupingVariables <- c('species','ageInSamples','year','season','age')
groupingVariablesNoYear <- c('species','ageInSamples','season','age')
riverGroupVar <- NULL

# Select THIS for metaPop analysis, or
#groupingVariables <- c('species','metaPop01','ageInSamples', 'year','season','age')
#groupingVariablesNoYear <- c('species','metaPop01','ageInSamples','season','age')
#riverGroupVar <- groupingVariables[2]

# THIS for 4 river analysis
#groupingVariables <- c('species','riverOrdered','ageInSamples','year','season','age')
#groupingVariablesNoYear <- c('species','riverOrdered','ageInSamples','season','age')
#riverGroupVar <- groupingVariables[2]
#####################################################
#####################################################

# Has mean size changed over time?
means <- 
  cd %>% mutate( age = year - cohort,
                 metaPop01 = ifelse(riverOrdered == 'wb obear', 0 ,1)) %>%
    filter( ageInSamples %in% 1:9, age < 3, year < 2016, area %in% c('inside','trib'), enc == 1
            #!(riverOrdered == 'wb mitchell' & ageInSamples > 8) #low sample sizes
            ) %>% 
    group_by_at(vars(one_of(groupingVariables))) %>%
    summarize(meanLen = mean(observedLength,na.rm = TRUE),
              sdLen = sd(observedLength,na.rm = TRUE),
              meanGR = mean(grLength,na.rm = TRUE),
              sdGR = sd(grLength,na.rm = TRUE),
              n = n()
              )

means$seasonGG <- factor(means$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
means$speciesGG <- factor(means$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
means$metaPop01GG <- factor(means$metaPop01, levels = c(0,1), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

lenSlopes <- means %>%
  group_by_at(vars(one_of(groupingVariablesNoYear))) %>%
  #group_by(species,metaPop01,ageInSamples,year,season,age) %>%
  nest()

lengthModel <- function(df){
  lm( meanLen ~ year, data = df )
}

grModel <- function(df){
  lm( meanGR ~ year, data = df )
}

lenSlopes <- lenSlopes %>%
  mutate( model = map(lenSlopes$data, lengthModel),
          modelGR = map(lenSlopes$data, grModel))

###################################
######### len here, gr below ######
lenSlopesTidy <- lenSlopes %>%
  mutate(glance = map(model, broom::tidy)) %>% 
  unnest(glance, .drop = TRUE) %>%
  filter(term == 'year') %>%
  mutate( sig = ifelse(p.value < 0.05,"*","")) %>%
  mutate( sig = ifelse(p.value < 0.001,"**",sig)) %>%
  mutate( sig = ifelse(p.value < 0.0001,"***",sig))  %>%
  filter(!is.nan(statistic))

lenSlopesTidy$seasonGG <- factor(lenSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
lenSlopesTidy$speciesGG <- factor(lenSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

save(lenSlopesTidy,file = paste0('D:/projects/westBrook-patterns/data/out/lenSlopesTidy_',riverGroupVar,'.RData'))
##

ggplot( means %>% filter(year >= 0), aes(year,meanLen, group= speciesGG)) +
  geom_point( aes(color = speciesGG), size = 2) +
  geom_smooth( aes(color = speciesGG), se = FALSE, method = 'lm') +
  labs( x = "Year", y = "Mean body length (mm)" ) +
  # geom_text(data = lenSlopesTidy %>% filter(species == 'bkt'), aes(x=1999,y=220, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
  # geom_text(data = lenSlopesTidy %>% filter(species == 'bnt'), aes(x=1999,y=200, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
  # geom_text(data = lenSlopesTidy %>% filter(species == 'ats'), aes(x=1999,y=180, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
  geom_text(data = lenSlopesTidy %>% filter(species == 'bkt', age<2), aes(x=2000,y=225, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
  geom_text(data = lenSlopesTidy %>% filter(species == 'bnt', age<2), aes(x=2000,y=203, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
#  geom_text(data = lenSlopesTidy %>% filter(species == 'ats', age<2), aes(x=2000,y=181, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +

    geom_text(data = lenSlopesTidy %>% filter(species == 'bkt', age==2), aes(x=2009,y=104, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
  geom_text(data = lenSlopesTidy %>% filter(species == 'bnt', age==2), aes(x=2009,y=82, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
#  geom_text(data = lenSlopesTidy %>% filter(species == 'ats', age==2), aes(x=2012.5,y=80, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
  theme_publication() +
  scale_color_grey(start = 0.15, end = 0.75) +
 # facet_grid(seasonGG ~ riverOrdered + age)
  facet_grid(as.formula(paste("seasonGG ~", riverGroupVar, "+ age")))

ggsave(paste0('figures/meanLengths_',riverGroupVar,'.png'), width = 11, height = 11, units='in')

# mean increases 
lenSlopesTidy %>%
  filter() %>%
  summarise( means = mean(estimate),
             sd = sd(estimate))

# mean increases by species
lenSlopesTidy %>%
  group_by(species) %>%
  summarise( means = mean(estimate),
             sd = sd(estimate))

lenSlopesTidy %>%
  group_by(species, metaPop01) %>%
  summarise( means = mean(estimate),
             sd = sd(estimate))

# do increases (slopes) increase over age? Yes
#ggplot(lenSlopesTidy %>% filter(!(riverOrdered == 'wb mitchell' & ageInSamples > 8)), aes((ageInSamples),estimate, color=speciesGG)) +
ggplot(lenSlopesTidy %>% filter(ageInSamples %in% 1:9), aes((ageInSamples),estimate, color=speciesGG)) +
  geom_point( size = 3) +
  geom_smooth(method='lm', se=FALSE) +
  scale_x_continuous(breaks = 1:10, labels = c(0,0,1,1,1,1,2,2,2,2)) +
  labs( x = "Age", y = "Slope of mean body length over years" ) +
  theme_publication() +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_wrap(as.formula(paste("~", riverGroupVar)))

ggsave(paste0('figures/meanLengthsOverAge_',riverGroupVar,'.png'), width = 6, height = 6, units='in')

# just bkt
ggplot(lenSlopesTidy %>% filter(ageInSamples %in% 1:9, species == 'bkt'), aes_(~ageInSamples, ~estimate, group = as.name(riverGroupVar))) +
  geom_point(aes_(color= as.name(riverGroupVar)), size = 3) +
  geom_smooth(aes_(color= as.name(riverGroupVar)), method='lm', se=FALSE) +
  scale_x_continuous(breaks = 1:10, labels = c(0,0,1,1,1,1,2,2,2,2)) +
  labs( x = "Age", y = "Slope of mean body length over years" ) +
  theme_publication()

ggsave(paste0('figures/meanLengthsOverAgeBKT_',riverGroupVar,'.png'), width = 6, height = 6, units='in')

summary(lm(estimate~ageInSamples, data = lenSlopesTidy %>% filter(species=='bkt')))
summary(lm(estimate~ageInSamples, data = lenSlopesTidy %>% filter(species=='bnt')))

############################################################################################################
#########################
#### gr here, len above

grSlopesTidy <- lenSlopes %>%
  mutate(glance = map(modelGR, broom::tidy)) %>% 
  unnest(glance, .drop = TRUE) %>%
  filter(term == 'year') %>%
  mutate( sig = ifelse(p.value < 0.05,"*","")) %>%
  mutate( sig = ifelse(p.value < 0.001,"**",sig)) %>%
  mutate( sig = ifelse(p.value < 0.0001,"***",sig))  %>%
  filter(!is.nan(statistic))

grSlopesTidy$seasonGG <- factor(grSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
grSlopesTidy$speciesGG <- factor(grSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

save(grSlopesTidy,file = paste0('D:/projects/westBrook-patterns/data/out/grSlopesTidy_',riverGroupVar,'.RData'))
##

##
means$seasonGG <- factor(means$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
means$speciesGG <- factor(means$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
grSlopesTidy$seasonGG <- factor(grSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
grSlopesTidy$speciesGG <- factor(grSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

ggplot( means %>% filter(), aes(year,meanGR, group= speciesGG)) +
  geom_point( aes(color = speciesGG), size = 4) +
  geom_smooth( aes(color = speciesGG), se = FALSE, method = 'lm') +
  labs( x = "Year", y = "Mean growth rate (mm/d)" ) +
  # geom_text(data = grSlopesTidy %>% filter(species == 'bkt'), aes(x=1999,y=220, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
  # geom_text(data = grSlopesTidy %>% filter(species == 'bnt'), aes(x=1999,y=200, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
  # geom_text(data = grSlopesTidy %>% filter(species == 'ats'), aes(x=1999,y=180, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
  geom_text(data = grSlopesTidy %>% filter(species == 'bkt', age<2), aes(x=2000,y=0.7, label=paste(round(estimate,4),sig)),hjust = 0)+#, color = speciesGG)) +
  geom_text(data = grSlopesTidy %>% filter(species == 'bnt', age<2), aes(x=2000,y=0.6, label=paste(round(estimate,4),sig)),hjust = 0)+#, color = speciesGG)) +
#  geom_text(data = grSlopesTidy %>% filter(species == 'ats', age<2), aes(x=2000,y=181, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +

    geom_text(data = grSlopesTidy %>% filter(species == 'bkt', age==2), aes(x=2011.5,y=0.66, label=paste(round(estimate,4),sig)),hjust = 0)+#, color = speciesGG)) +
  geom_text(data = grSlopesTidy %>% filter(species == 'bnt', age==2), aes(x=2011.5,y=0.56, label=paste(round(estimate,4),sig)),hjust = 0)+#, color = speciesGG)) +
#  geom_text(data = grSlopesTidy %>% filter(species == 'ats', age==2), aes(x=2012.5,y=80, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
  theme_publication() +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_grid(as.formula(paste("seasonGG ~", riverGroupVar, "+ age")))

ggsave(paste0('figures/meanGRs_',riverGroupVar,'.png'), width = 11, height = 11, units='in')

# mean increases 
grSlopesTidy %>%
  filter() %>%
  summarise( means = mean(estimate),
             sd = sd(estimate))

# mean increases by species
grSlopesTidy %>%
  group_by(species) %>%
  summarise( means = mean(estimate),
             sd = sd(estimate))

# do increases (slopes) increase over age? Yes
#ggplot(lenSlopesTidy %>% filter(!(riverOrdered == 'wb mitchell' & ageInSamples > 8)), aes((ageInSamples),estimate, color=speciesGG)) +
ggplot(grSlopesTidy %>% filter(ageInSamples %in% 1:9), aes((ageInSamples),estimate, color=speciesGG)) +
  geom_point( size = 3) +
  geom_smooth(method='lm', se=FALSE) +
  scale_x_continuous(breaks = 1:10, labels = c(0,0,1,1,1,1,2,2,2,2)) +
  labs( x = "Age", y = "Slope of mean body length over years" ) +
  theme_publication() +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_wrap(as.formula(paste("~", riverGroupVar)))

ggsave(paste0('figures/meanGRsOverAge_',riverGroupVar,'.png'), width = 6, height = 6, units='in')

# just bkt
ggplot(grSlopesTidy %>% filter(ageInSamples %in% 1:9, species == 'bkt'), aes_(~ageInSamples, ~estimate, group = as.name(riverGroupVar))) +
  geom_point(aes_(color= as.name(riverGroupVar)), size = 3) +
  geom_smooth(aes_(color= as.name(riverGroupVar)), method='lm', se=FALSE) +
  scale_x_continuous(breaks = 1:10, labels = c(0,0,1,1,1,1,2,2,2,2)) +
  labs( x = "Age", y = "Slope of mean body length over years" ) +
  theme_publication()

ggsave(paste0('figures/meanGRsOverAgeBKT_',riverGroupVar,'.png'), width = 6, height = 6, units='in')

summary(lm(estimate~ageInSamples, data = grSlopesTidy %>% filter(species=='bkt')))
summary(lm(estimate~ageInSamples, data = grSlopesTidy %>% filter(species=='bnt')))

```{r where are the fish?}

allFishBySpeciesYOY$seasonGG <- factor(allFishBySpeciesYOY$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)

allFishBySpeciesYOY$speciesGG <- factor(allFishBySpeciesYOY$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

allFishBySpeciesYOY$riverGG <- factor(allFishBySpeciesYOY$river,
                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
                         labels = c("WB","OL","OS","IL"), ordered = T)

>>>>

---
title: "Body size patterns"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

@@

```{r gams for body size - model selection}
speciesIn <- 'bnt'
depVarIn <- 'observedLength'
kIn <- 4 # # of knots for the speces and river smooths

@@

vB$riverFactor <- factor(vB$river, levels=c('west brook', 'wb jimmy', 'wb mitchell', 'wb obear'))
#vB$riverOrdered <- riverOrdered exists

@@

vB$tagIndexFactor <- factor(vB$tagIndex)

#vB$riverOrdered <- factor(vB$river,
#                         levels=c('west brook', 'wb jimmy', 'wb mitchell', 'wb obear'),
#                         ordered = T)

dataIn <- vB %>% filter(cohort %in% 2000:2013, species %in% speciesIn)
kIn <- 4
################################################################################

@@

mod <- bam(observedLength ~ cohort * river + 
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")
summary(mod)

@@

##########################################################
# by cohortSpeciesRiver

system.time(
modCR <- bam(observedLength ~ cohort * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered), 
             data = dataIn, method = "REML")
)
summary(modCR)
#plot(modCR, pages = 8, scale = 0)

@@

ggplot(predGrid2CR, aes(ydayCumul, predCR)) +
  geom_point() + 
  geom_line() +
  ylim(50,210) +
  facet_grid(river ~ cohort)+
  ggtitle('modCR')

##########################################################
# by cohortSpeciesRiver with s(cohort)

system.time(
modCR_c <- bam(observedLength ~ cohort * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn), 
             data = dataIn, method = "REML")
)
summary(modCR_c)
#plot(modCR_c, pages = 8, scale = 0)

predGridCR_c <- expand.grid(cohortRiverOrdered = unique(modCR$model$cohortRiverOrdered),
                           ydayCumul = seq(250,1000, by = 30)) %>%
                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
                           riverOrdered = substring(cohortRiverOrdered,6),
                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
                           cohortCentered = cohortOrdered - cohortOffset,
                           cohort = cohortOrdered,
                           river = riverOrdered)

predCR_c <- predict(modCR_c, newdata = predGridCR_c)#, type = 'lpmatrix')

predGrid2CR_c <- bind_cols(predGridCR_c, as.data.frame(predCR_c))

ggplot(predGrid2CR_c, aes(ydayCumul, predCR_c, group = cohort, color = cohort)) +
  geom_point() + 
  geom_line() +
  #ylim(50,210) +
  facet_wrap(~river)+
  ggtitle('modCR_c')

##########################################################
# by cohortRiver - w ind RE

# need more memory
# system.time(
# modCR_Ind <- bam(observedLength ~ cohort * river +
#                s(ydayCumul, bs = 'cr') +
#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
#                s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
#                s(ydayCumul, tagIndexFactor, bs="fs", m=1, k = kIn),
#              data = dataIn, method = "REML")
# )
# summary(modCR_Ind)
# plot(modCR_Ind, pages = 8, scale = 0)

##########################################################
# by cohortSpeciesRiver, additive in s()

system.time(
modC_R <- bam(observedLength ~ cohort * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered), 
             data = dataIn, method = "REML")
)
summary(modC_R)
#plot(modC_R, pages = 4, scale = 0)

@@

##########################

modC <- bam(observedLength ~ cohort + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered), 
             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt')), method = "REML")

modR <- bam(observedLength ~ river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered), 
             data = dataIn, method = "REML")

AIC(mod,modC,modR,modC_R,modCR,modCR_c) %>% rownames_to_column() %>% arrange(AIC)
save(mod,modC,modR,modC_R,modCR,modCR_c, file = paste0('./data/out/mods_',speciesIn,'.RData'))
######

@@

```{r best models}

# Brook trout

modBKT <- bam(observedLength ~ cohort * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn), 
             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bkt'), method = "REML")

summary(modBKT)
#plot(modBKT, pages = 8, scale = 0)

predGridBKT <- expand.grid(cohortRiverOrdered = unique(modBKT$model$cohortRiverOrdered),
                           ydayCumul = seq(250,1000, by = 30)) %>%
                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
                           riverOrdered = substring(cohortRiverOrdered,6),
                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
                           cohortCentered = cohortOrdered - cohortOffset,
                           cohort = cohortOrdered,
                           river = riverOrdered)

predBKT <- predict(modBKT, newdata = predGridBKT)#, type = 'lpmatrix')

predGrid2BKT <- bind_cols(predGridBKT, as.data.frame(predBKT)) %>% mutate(pred = predBKT, species = 'bkt')

#################################
# Brown trout

modBNT <- bam(observedLength ~ cohort * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn), 
             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bkt'), method = "REML")

summary(modBNT)
#plot(modBNT, pages = 8, scale = 0)

predGridBNT <- expand.grid(cohortRiverOrdered = unique(modBNT$model$cohortRiverOrdered),
                           ydayCumul = seq(250,1000, by = 30)) %>%
                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
                           riverOrdered = substring(cohortRiverOrdered,6),
                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
                           cohortCentered = cohortOrdered - cohortOffset,
                           cohort = cohortOrdered,
                           river = riverOrdered)

predBNT <- predict(modBNT, newdata = predGridBNT)#, type = 'lpmatrix')

predGrid2BNT <- bind_cols(predGridBNT, as.data.frame(predBNT)) %>% mutate(pred = predBNT, species = 'bnt')

##################
# combine species' predictions

predGrid2Both <- bind_rows(predGrid2BKT, predGrid2BNT) %>% select(-predBKT, -predBNT)

predGrid2Both$riverGG <- factor(predGrid2Both$riverOrdered,
                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
                         labels = c("West Brook","Open Large","Open Small","Isolated Small"), ordered = T)
predGrid2Both$speciesGG <- factor(predGrid2Both$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

ggplot(predGrid2Both, aes(ydayCumul, pred, group = cohort, color = cohort)) +
  geom_point() + 
  geom_line() +
  labs(y = 'Predicted body size (mm)',
       x = 'Days since age-0 Jan 1',
       color = 'Cohort') +
  scale_color_viridis_c(breaks = c(2000,2004,2008,2012), labels = c(2000,2004,2008,2012)) +
  theme_publication() +
  theme(legend.position = "right", 
        legend.direction = "vertical",
        legend.title = element_text(face="plain")) +
  guides(color = guide_colorbar(barheight = 10,
                                barwidth = 1.5,
                                ticks = FALSE,
                                nbin = 100)) +
#  guide_legend(title = 'Cohort') +
  facet_grid(speciesGG ~ riverGG)

@@
